Signed Peripherality Functions for Multivariate Statistical Inference

### 2018-03-23 MS-new/1_Introduction.tex
main paper v1 done, proofs and intro remain

Consider a point $\bfx$ from some $\cX \subseteq \BR^p$, and the spatial sign function $\bfS : \cX \times \BR^p \rightarrow \BR^p$, defined as
%
$$
\bfS (\bfx; \bfmu_x) = \frac{\bfx - \bfmu_x}{\| \bfx - \bfmu_x \|}
\BI ( \bfx \ne \bfmu_x ),
$$
%
where $\bfmu_x \in \BR^p$ is a fixed location parameter, $\BI(\cdot)$ is the 0/1 indicator function and $\|.\|$ is the euclidean norm. This is a generalization of the real-valued case of the indicator for whether a point $x \in \BR$ is to the right, left or at a scalar $\mu_{x} \in \BR$, and was first introduced by \cite{MottonenOja95}. Given a data matrix $\bfX = (\bfx_{1}, \ldots, \bfx_{n})^T$, a popular approach of using spatial signs in the literature has been to take spatial signs of the data points, i.e. $\bfS(\bfx_i)$, and use the matrix of signs as a proxy of $\bfX$ to formulate robust versions of location and scale problems  \citep{locantore99, OjaBook10,WangPengLi15}.

@@

\begin{Definition}\label{defn:peripherality}
For a set of probability measures $\cM$ taking values in $\BR^p$, a {\it peripherality function} $ P : \BR^p \times \cM \rightarrow \BR^+ \cup \{ 0\}$,  is a function that satisfies the following condition:\\

@@

The notion of multivariate ranks goes back to \cite{PuriSenBook}, where they take the vector consisting of marginal univariate ranks as multivariate rank vector. Subsequent definitions of multivariate ranks were proposed by \cite{MottonenOja95,HallinPaindaveine02} and \cite{Chernozhukov14}. Compared to them, the advantages of our signed peripherality function are two-fold: interpretability and flexibility. Firstly, as evident in Figure~\ref{fig:rankplot}, our rank transformation preserves the general shape of the data. In fact, if the map $\| \bfx \| \mapsto P(\bfx, \BF)$ is one-one (given a fixed $\BF$), then it is possible to get back $\bfx$ from $\bfR (\bfx; \bfmu_x, \BF)$. It also provides an intuitive extension to any spatial sign-based methodology. Secondly, there is considerable scope of generalization within the present framework. Given any general inner product space, the definition of signed peripherality remains the same when euclidean norm is replaced by the norm induced by the inner product. Even within $\BR^p$, it is possible to generalize the ranking function in \eqref{eqn:rank} by adding separate transformations on the sign and peripherality functions to obtain alternate representations of the data in lower or higher dimensional spaces. However, we limit our discussion in this paper to the original formulation in \eqref{eqn:rank}.

>>>>

Consider a point $\bfx$ from some $\cX \subseteq \BR^p$, and the spatial sign function $\bfS : \cX \times \BR^p \mapsto \BR^p$, defined as
%
$$
\bfS (\bfx; \bfmu_x) = \frac{\bfx - \bfmu_x}{\| \bfx - \bfmu_x \|}
\BI ( \bfx \ne \bfmu_x ),
$$
%
where $\bfmu_x \in \BR^p$ is a fixed location parameter, $\BI(\cdot)$ is the 0/1 indicator function and $\|.\|$ is the euclidean norm. This is a generalization of the real-valued case of the indicator for whether a point $x \in \BR$ is to the right, left or at a scalar $\mu_{x} \in \BR$, and was first introduced by \cite{MottonenOja95}. Given a data matrix $\bfX = (\bfx_{1}, \ldots, \bfx_{n})^T$, a popular approach of using spatial signs in the literature has been to take spatial signs of the data points, i.e. $\bfS(\bfx_i)$, and use the matrix of signs as a proxy of $\bfX$ to formulate robust versions of location and scale problems  \citep{locantore99, OjaBook10,WangPengLi15}.

@@

\begin{Definition}\label{defn:peripherality}
For a set of probability measures $\cM$ taking values in $\BR^p$, a {\it peripherality function} $ P : \BR^p \times \cM \mapsto \BR^+ \cup \{ 0\}$,  is a function that satisfies the following condition:\\

@@

The notion of multivariate ranks goes back to \cite{PuriSenBook}, where they take the vector consisting of marginal univariate ranks as multivariate rank vector. Subsequent definitions of multivariate ranks were proposed by \cite{MottonenOja95,HallinPaindaveine02} and \cite{Chernozhukov14}. Compared to them, the advantages of our signed peripherality function are two-fold: interpretability and flexibility. Firstly, as evident in Figure~\ref{fig:rankplot}, our rank transformation preserves the general shape of the data. In fact, if the map $\| \bfx \| \rightarrow P(\bfx, \BF)$ is one-one (given a fixed $\BF$), then it is possible to get back $\bfx$ from $\bfR (\bfx; \bfmu_x, \BF)$. It also provides an intuitive extension to any spatial sign-based methodology. Secondly, there is considerable scope of generalization within the present framework. Given any general inner product space, the definition of signed peripherality remains the same when euclidean norm is replaced by the norm induced by the inner product. Even within $\BR^p$, it is possible to generalize the ranking function in \eqref{eqn:rank} by adding separate transformations on the sign and peripherality functions to obtain alternate representations of the data in lower or higher dimensional spaces. However, we limit our discussion in this paper to the original formulation in \eqref{eqn:rank}.
